---
title: "Logistische Regression zur Brustkrebs-Diagnose mit Feature Selection"
author: "Data Analysis"
date: "`r Sys.Date()`"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.width = 10, fig.height = 6, dpi = 300)

# Set seed for reproducibility
set.seed(123)

# Check and install missing packages
required_packages <- c(
  "readxl", "dplyr", "tidyr", "stringr",
  "ggplot2", "broom", "car",
  "glmnet", "pROC", "forcats", "corrplot", "stargazer"
)

# Function to check and install packages
install_if_missing <- function(pkg) {
  if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
    cat("Versuche Paket zu installieren:", pkg, "\n")
    install.packages(pkg, repos = "http://cran.rstudio.com/", dependencies = TRUE)
    if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
      stop("Fehler: Paket '", pkg, "' konnte nicht geladen werden.\n",
           "Bitte installieren Sie es manuell in R:\n",
           "  install.packages('", pkg, "', dependencies = TRUE)\n",
           "Falls das Problem mit 'car' auftritt, versuchen Sie:\n",
           "  install.packages(c('nloptr', 'carData', 'pbkrtest', 'lme4', 'car'), dependencies = TRUE)")
    }
  }
}

# Install and load packages one by one
cat("Prüfe und installiere Pakete bei Bedarf...\n")
for(pkg in required_packages) {
  install_if_missing(pkg)
}
cat("Alle Pakete erfolgreich geladen.\n")
```

# 1. Setup

Dieses Dokument führt eine logistische Regression zur Vorhersage der Brustkrebs-Diagnose durch. Der Fokus liegt auf einer robusten Feature Selection unter Vermeidung von Multikollinearität.

**Zielvariable:** diagnosis (M = Malignant = 1, B = Benign = 0)

**Methodik:**
- Korrelationsfilter (|r| > 0.9)
- VIF-Check (VIF < 5)
- LASSO-Regularisierung für Feature Selection
- Finales GLM-Modell auf selektierten Features

```{r load-data, echo=TRUE}
# Read Excel file
data_raw <- read_excel("dataset_stat.xlsx", skip = 1)

# Check structure
cat("Original dataset dimensions:", dim(data_raw), "\n")
cat("Column names (first 10):", paste(head(colnames(data_raw), 10), collapse = ", "), "\n")

# Remove header row if it contains variable names
if(nrow(data_raw) > 0 && !is.na(data_raw[1, 1]) && data_raw[1, 1] == "id") {
  data_raw <- data_raw[-1, ]
}

cat("After cleaning, dimensions:", dim(data_raw), "\n")
path <- paste(getwd(), "/dataset_stat.xlsx", sep="")
cat(path)
file.exists(path)
View(data_raw)

```

# 2. Data Load & Cleaning

```{r preprocessing, echo=TRUE}
# Identify ID column (common names: id, ID, X, or first column if it looks like ID)
id_cols <- which(tolower(colnames(data_raw)) %in% c("id", "x") | 
                 grepl("^id", tolower(colnames(data_raw)), ignore.case = TRUE))
if(length(id_cols) > 0) {
  id_col_name <- colnames(data_raw)[id_cols[1]]
  cat("ID column identified:", id_col_name, "\n")
  data <- data_raw %>% select(-all_of(id_col_name))
} else {
  # Check if first column looks like ID (all unique values)
  if(length(unique(data_raw[[1]])) == nrow(data_raw)) {
    cat("First column appears to be ID (all unique), removing it\n")
    data <- data_raw %>% select(-1)
  } else {
    data <- data_raw
    cat("No ID column identified, keeping all columns\n")
  }
}

# Convert diagnosis to binary: M = 1, B = 0
data$diagnosis <- as.character(data$diagnosis)
data$diagnosis_binary <- ifelse(data$diagnosis == "M", 1, 0)

cat("\nDiagnosis distribution:\n")
print(table(data$diagnosis, useNA = "ifany"))
cat("\nBinary diagnosis distribution:\n")
print(table(data$diagnosis_binary, useNA = "ifany"))

# Separate predictors and outcome
predictors_data <- data %>% 
  select(-diagnosis, -diagnosis_binary)

# Convert all columns to numeric (they should be numeric but might be character)
predictors_data <- predictors_data %>%
  mutate_all(~as.numeric(as.character(.))) %>%
  select_if(~!all(is.na(.)))  # Remove columns that are all NA

# Check for missing values
missing_counts <- colSums(is.na(predictors_data))
if(sum(missing_counts) > 0) {
  cat("\nMissing values found:\n")
  print(missing_counts[missing_counts > 0])
  # Remove rows with missing values
  complete_rows <- complete.cases(predictors_data)
  predictors_data <- predictors_data[complete_rows, ]
  data <- data[complete_rows, ]
  cat("\nRemoved", sum(!complete_rows), "rows with missing values\n")
} else {
  cat("\nNo missing values found in predictors\n")
}

# Final dataset
y <- data$diagnosis_binary
X <- predictors_data

cat("\nFinal dataset dimensions:", dim(X), "\n")
cat("Number of predictors:", ncol(X), "\n")
cat("Number of observations:", nrow(X), "\n")
```

# 3. Explorative Checks

```{r exploratory, echo=TRUE}
# Class balance
cat("Class balance:\n")
class_table <- table(y)
print(class_table)
cat("Proportions:\n")
print(prop.table(class_table))

# Summary statistics of predictors
if(ncol(X) > 0) {
  cat("\nSummary of predictors (first 5):\n")
  print(summary(X[, 1:min(5, ncol(X))]))
} else {
  cat("\nWarnung: Keine Prädiktoren verfügbar!\n")
}
```

```{r class-distribution-plot, echo=TRUE, fig.width=8, fig.height=6}
# Barplot: Verteilung der Zielvariable (Klassen M und B)
diagnosis_dist <- data.frame(
  Diagnosis = c("Benign (B)", "Malignant (M)"),
  Count = c(sum(data$diagnosis == "B"), sum(data$diagnosis == "M"))
)

class_distribution_plot <- ggplot(diagnosis_dist, aes(x = Diagnosis, y = Count, fill = Diagnosis)) +
  geom_bar(stat = "identity", alpha = 0.8) +
  scale_fill_manual(values = c("Benign (B)" = "#2E8B57", "Malignant (M)" = "#CD5C5C"),
                    guide = "none") +
  geom_text(aes(label = Count), vjust = -0.3, size = 5, fontface = "bold") +
  labs(
    title = "Distribution of Diagnosis Classes",
    subtitle = "Number of Benign (B) vs. Malignant (M) Cases",
    x = "Diagnosis",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  ylim(0, max(diagnosis_dist$Count) * 1.1)

print(class_distribution_plot)

# Save plot
ggsave("class_distribution.png", class_distribution_plot, width = 8, height = 6, dpi = 300)
cat("\nClass distribution plot saved as: class_distribution.png\n")
```

# 4. Multicollinearity Handling

## 4.1 Korrelationsfilter

```{r correlation-filter, echo=TRUE}
# Calculate correlation matrix
cor_matrix <- cor(X, use = "complete.obs")

# Find highly correlated pairs (|r| > 0.9)
high_cor <- which(abs(cor_matrix) > 0.9 & cor_matrix != 1, arr.ind = TRUE)

if(nrow(high_cor) > 0) {
  # Create a list of variables to remove (keep one from each pair)
  vars_to_remove <- c()
  for(i in 1:nrow(high_cor)) {
    var1 <- rownames(cor_matrix)[high_cor[i, 1]]
    var2 <- colnames(cor_matrix)[high_cor[i, 2]]
    if(!var1 %in% vars_to_remove && !var2 %in% vars_to_remove) {
      # Remove the second one (arbitrary choice)
      vars_to_remove <- c(vars_to_remove, var2)
    }
  }
  
  cat("Highly correlated variables (|r| > 0.9) found. Removing:", length(vars_to_remove), "variables\n")
  print(vars_to_remove)
  
  X <- X %>% select(-all_of(vars_to_remove))
  cat("\nAfter correlation filter, remaining predictors:", ncol(X), "\n")
} else {
  cat("No highly correlated pairs (|r| > 0.9) found\n")
}

# Correlation heatmap (if not too many variables)
if(ncol(X) <= 30) {
  # Erstelle Korrelationsmatrix-Plot
  cor_matrix_plot <- cor(X, use = "complete.obs")
  
  # Plot in Konsole anzeigen
  corrplot(cor_matrix_plot, type = "upper", 
           order = "hclust", tl.cex = 0.7, tl.col = "black")
  
  # Speichere Korrelationsmatrix als PNG
  png("correlation_matrix.png", width = 12, height = 10, units = "in", res = 300)
  corrplot(cor_matrix_plot, type = "upper", 
           order = "hclust", tl.cex = 0.7, tl.col = "black",
           title = "Correlation Matrix", mar = c(0, 0, 2, 0))
  dev.off()
  
  cat("\nKorrelationsmatrix wurde gespeichert als: correlation_matrix.png\n")
} else {
  cat("\nZu viele Variablen (", ncol(X), ") für Korrelationsmatrix-Plot. Überspringe Visualisierung.\n")
}
```

## 4.2 VIF-Check und iterative Bereinigung

```{r vif-check, echo=TRUE}
# Create initial model for VIF calculation
# We need to fit a GLM first to calculate VIF
initial_model <- glm(y ~ ., data = data.frame(y = y, X), family = binomial)

# Calculate VIF
vif_values <- vif(initial_model)

cat("Initial VIF values:\n")
print(sort(vif_values, decreasing = TRUE))

# Iteratively remove variables with VIF >= 5
vif_threshold <- 5
X_vif_cleaned <- X
max_iterations <- 50
iteration <- 0

while(iteration < max_iterations) {
  # Fit model
  temp_model <- glm(y ~ ., data = data.frame(y = y, X_vif_cleaned), 
                    family = binomial, maxit = 100)
  
  # Check VIF
  vif_current <- tryCatch(vif(temp_model), error = function(e) NULL)
  
  if(is.null(vif_current)) {
    cat("VIF calculation failed, breaking loop\n")
    break
  }
  
  max_vif <- max(vif_current)
  
  if(max_vif < vif_threshold) {
    cat("\nAll VIF values <", vif_threshold, "after", iteration, "iterations\n")
    break
  }
  
  # Remove variable with highest VIF
  var_to_remove <- names(which.max(vif_current))
  
  # Check if variable exists in data frame
  if(!var_to_remove %in% colnames(X_vif_cleaned)) {
    cat("Warning: Variable", var_to_remove, "not found in data, breaking loop\n")
    break
  }
  
  cat("Iteration", iteration + 1, ": Removing", var_to_remove, 
      "(VIF =", round(max_vif, 2), ")\n")
  
  X_vif_cleaned <- X_vif_cleaned %>% select(-all_of(var_to_remove))
  iteration <- iteration + 1
}

# Final VIF check
final_model_vif <- glm(y ~ ., data = data.frame(y = y, X_vif_cleaned), 
                       family = binomial, maxit = 100)
vif_final <- vif(final_model_vif)

cat("\nFinal VIF values (all <", vif_threshold, "):\n")
print(sort(vif_final, decreasing = TRUE))

cat("\nVariables remaining after VIF filtering:", ncol(X_vif_cleaned), "\n")

# Create VIF comparison table
vif_comparison <- data.frame(
  Variable = names(vif_final),
  VIF_Final = round(vif_final, 3)
) %>%
  arrange(desc(VIF_Final))

print(vif_comparison)
```

# 5. Feature Selection mit LASSO

```{r lasso-selection, echo=TRUE}
# Prepare data for glmnet (requires matrix)
X_matrix <- as.matrix(X_vif_cleaned)

# Standardize predictors for LASSO
X_scaled <- scale(X_matrix)
X_scaled[is.nan(X_scaled)] <- 0  # Handle any NaN from scaling

# Cross-validated LASSO
cv_lasso <- cv.glmnet(X_scaled, y, family = "binomial", alpha = 1, 
                      nfolds = 10, type.measure = "auc")

# Plot cross-validation results
plot(cv_lasso)

# Get lambda.min and lambda.1se
lambda_min <- cv_lasso$lambda.min
lambda_1se <- cv_lasso$lambda.1se

cat("Lambda min:", lambda_min, "\n")
cat("Lambda 1se:", lambda_1se, "\n")

# Use lambda.min for more features, lambda.1se for more parsimony
lambda_selected <- lambda_min
cat("\nUsing lambda.min for feature selection\n")

# Get coefficients at selected lambda
lasso_coef <- coef(cv_lasso, s = lambda_selected)
lasso_features <- rownames(lasso_coef)[which(lasso_coef != 0)][-1]  # Exclude intercept

cat("\nSelected features by LASSO (", length(lasso_features), "features):\n")
print(lasso_features)

# Selected features dataframe
selected_features_df <- data.frame(
  Feature = lasso_features,
  Method = "LASSO"
)

# Save selected features
write.csv(selected_features_df, "selected_features.csv", row.names = FALSE)
```

# 6. Finale Logistische Regression

```{r final-glm, echo=TRUE}
# Refit GLM on selected features (non-scaled data for interpretable coefficients)
X_final <- X_vif_cleaned %>% select(all_of(lasso_features))

# Fit final model
final_model <- glm(y ~ ., data = data.frame(y = y, X_final), 
                   family = binomial(link = "logit"))

# Model summary
cat("Final Model Summary:\n")
print(summary(final_model))

# Extract coefficients #Finale Koeffizienten-Summary 
coef_summary <- broom::tidy(final_model, conf.int = TRUE, exponentiate = FALSE)

# Calculate Odds Ratios and CI, Erweiterung um Odds Ratios
coef_summary$odds_ratio <- exp(coef_summary$estimate)
coef_summary$odds_ratio_ci_lower <- exp(coef_summary$conf.low)
coef_summary$odds_ratio_ci_upper <- exp(coef_summary$conf.high)

# Sort by p-value (most significant first)
coef_summary <- coef_summary %>% 
  arrange(p.value) %>%
  mutate(
    p_value_formatted = ifelse(p.value < 0.001, "< 0.001", 
                               sprintf("%.4f", p.value)),
    significant = ifelse(p.value < 0.05, "Yes", "No")
  )

cat("\nFinal Coefficients (sorted by p-value):\n")
#finale summary
print(coef_summary %>% 
      select(term, estimate, std.error, statistic, p.value, odds_ratio,
             odds_ratio_ci_lower, odds_ratio_ci_upper, significant))

# Save coefficients table
write.csv(coef_summary, "final_glm_coefficients.csv", row.names = FALSE)

# Reskalierung für bessere Interpretierbarkeit der Odds Ratios
# HINWEIS: Dies ist KEINE Standardisierung, sondern ein Rescaling nur aus Gründen 
# der besseren Lesbarkeit der Odds Ratios. Variablen mit sehr kleiner Skalierung 
# werden multipliziert, damit die Odds Ratios interpretierbar werden, ohne die 
# statistischen Eigenschaften des Modells zu ändern.

# Identifiziere alle Variablen mit sehr kleiner Skalierung
# 1. Alle _se-Variablen (smoothness_se, concavity_se, compactness_se, symmetry_se)
vars_se <- grep("_se$", colnames(X_final), value = TRUE)

# 2. Spezifische _worst-Variablen mit sehr kleiner Skala
# (smoothness_worst, concave.points_worst, symmetry_worst)
vars_worst_small <- c("smoothness_worst", "concave.points_worst", "concave points_worst", "symmetry_worst")
vars_worst_small <- vars_worst_small[vars_worst_small %in% colnames(X_final)]

# Kombiniere alle zu reskalierenden Variablen
vars_to_rescale <- unique(c(vars_se, vars_worst_small))

if(length(vars_to_rescale) > 0) {
  cat("\nReskaliere Variablen mit kleiner Skalierung (x1000):\n")
  cat("_se-Variablen:", paste(vars_se, collapse = ", "), "\n")
  if(length(vars_worst_small) > 0) {
    cat("_worst-Variablen:", paste(vars_worst_small, collapse = ", "), "\n")
  }
  cat("Gesamt:", length(vars_to_rescale), "Variablen\n")
  
  # Erstelle reskalierte Version von X_final
  X_final_rescaled <- X_final
  
  # Reskaliere identifizierte Variablen (Multiplikation mit 1000)
  for(var in vars_to_rescale) {
    X_final_rescaled[[var]] <- X_final_rescaled[[var]] * 1000
  }
  
  # Fitte Modell mit vollständig reskalierten Variablen
  final_model_rescaled <- glm(y ~ ., data = data.frame(y = y, X_final_rescaled), 
                              family = binomial(link = "logit"))
  
  # Berechne Odds Ratios des reskalierten Modells
  odds_ratios_rescaled <- exp(coef(final_model_rescaled))
  
  # Erstelle HTML-Tabelle mit stargazer (Odds Ratios anstelle von Logit-Koeffizienten)
  # Fußnote: Explizit erwähnen, welche Variablen reskaliert wurden
  rescale_note <- paste0(
    "Odds Ratios (exponentiated coefficients) are shown. ",
    "Selected variables with very small measurement scales were rescaled (x1000) for better interpretability: ",
    "all variables ending with '_se' (smoothness_se, concavity_se, compactness_se, symmetry_se) ",
    "and specific _worst variables (smoothness_worst, concave.points_worst, symmetry_worst). ",
    "Standard errors and p-values refer to the original logit coefficients. ",
    "This rescaling is for presentation purposes only and does not constitute standardization."
  )
  
  stargazer(final_model_rescaled, 
            type = "html",
            out = "final_glm_odds_ratios.htm",
            coef = list(odds_ratios_rescaled),
            p.auto = FALSE,
            title = "Final Logistic Regression Model: Odds Ratios",
            notes = rescale_note,
            notes.append = FALSE)
  
  cat("\nOdds Ratios Tabelle (mit vollständig reskalierten Variablen) wurde gespeichert als: final_glm_odds_ratios.htm\n")
} else {
  # Falls keine Variablen zum Reskalieren vorhanden, verwende ursprüngliches Modell
  cat("\nKeine Variablen mit kleiner Skalierung zum Reskalieren gefunden. Verwende ursprüngliches Modell.\n")
  odds_ratios <- exp(coef(final_model))
  
  # Setze final_model_rescaled und X_final_rescaled auf ursprüngliche Werte
  # damit sie für Forest Plot und Boxplots verfügbar sind
  final_model_rescaled <- final_model
  X_final_rescaled <- X_final
  
  stargazer(final_model, 
            type = "html",
            out = "final_glm_odds_ratios.htm",
            coef = list(odds_ratios),
            p.auto = FALSE,
            title = "Final Logistic Regression Model: Odds Ratios",
            notes = "Odds Ratios (exponentiated coefficients) are shown. Standard errors and p-values refer to the original logit coefficients.",
            notes.append = FALSE)
  
  cat("\nOdds Ratios Tabelle wurde gespeichert als: final_glm_odds_ratios.htm\n")
}

# Model fit statistics
cat("\nModel Fit Statistics:\n")
cat("AIC:", AIC(final_model), "\n")
cat("Null Deviance:", final_model$null.deviance, "\n")
cat("Residual Deviance:", final_model$deviance, "\n")
pseudo_r2 <- 1 - (final_model$deviance / final_model$null.deviance)
cat("Pseudo R-squared (McFadden):", round(pseudo_r2, 4), "\n")
```

**Hinweis:** "Höchster p-Wert" bedeutet üblicherweise die am wenigsten signifikanten Variablen. In diesem Report werden Features nach **niedrigstem p-Wert** sortiert, d.h. die signifikantesten Features erscheinen zuerst.

# 7. Visualisierungen

## 7.1 Forest Plot: Odds Ratios (Top 10 Features)

```{r forest-plot, echo=TRUE, fig.width=10, fig.height=7}
# Forest Plot basiert auf dem reskalierten Modell
# HINWEIS: Dieser Forest Plot verwendet das reskalierte Modell (final_model_rescaled),
# damit alle Odds Ratios sich auf sinnvolle, vergleichbare Einheitenänderungen beziehen.
# Reskalierte Prädiktoren (z.B. _se- und _worst-Variablen x1000) ermöglichen eine
# realistische Darstellung der Effekte pro Einheitserhöhung.

# Berechne coef_summary aus dem reskalierten Modell
coef_summary_rescaled <- broom::tidy(final_model_rescaled, conf.int = TRUE, exponentiate = FALSE)

# Berechne Odds Ratios und 95%-Konfidenzintervalle explizit aus dem reskalierten Modell
coef_summary_rescaled$odds_ratio <- exp(coef_summary_rescaled$estimate)
coef_summary_rescaled$odds_ratio_ci_lower <- exp(coef_summary_rescaled$conf.low)
coef_summary_rescaled$odds_ratio_ci_upper <- exp(coef_summary_rescaled$conf.high)

# Sortiere nach p-Wert und markiere Signifikanz
coef_summary_rescaled <- coef_summary_rescaled %>%
  arrange(p.value) %>%
  mutate(
    significant = ifelse(p.value < 0.05, "Yes", "No")
  )

# Select top 10 features (excluding intercept) aus dem reskalierten Modell
top10 <- coef_summary_rescaled %>%
  filter(term != "(Intercept)") %>%
  slice_head(n = 10) %>%
  mutate(term = fct_reorder(term, -p.value))  # Order by significance

# Create forest plot
forest_plot <- ggplot(top10, aes(x = odds_ratio, y = term)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point(aes(color = significant), size = 3) +
  geom_errorbarh(aes(xmin = odds_ratio_ci_lower, xmax = odds_ratio_ci_upper,
                     color = significant), height = 0.2) +
  scale_x_log10() +
  scale_color_manual(values = c("Yes" = "red", "No" = "blue"),
                     labels = c("Yes" = "Significant (p < 0.05)", 
                               "No" = "Not significant")) +
  labs(
    title = "Forest Plot: Odds Ratios (Top 10 Features)",
    subtitle = "95% Confidence Intervals on log scale (based on rescaled predictors)",
    x = "Odds Ratio (log scale)",
    y = "Feature",
    color = "Significance"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "bottom"
  )

print(forest_plot)

# Save plot
ggsave("top10_or_forest.png", forest_plot, width = 10, height = 7, dpi = 300)
```

## 7.2 Coefficient Plot: Beta Estimates (Top 10 Features)

```{r coefficient-plot, echo=TRUE, fig.width=10, fig.height=7}
# Coefficient plot
coef_plot <- ggplot(top10, aes(x = estimate, y = term)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_bar(aes(fill = estimate > 0), stat = "identity", width = 0.6) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, 
                 color = "black", alpha = 0.7) +
  scale_fill_manual(values = c("TRUE" = "#2E8B57", "FALSE" = "#CD5C5C"),
                    labels = c("TRUE" = "Positive", "FALSE" = "Negative"),
                    guide = guide_legend(title = "Direction")) +
  labs(
    title = "Coefficient Plot: Beta Estimates (Top 10 Features)",
    subtitle = "95% Confidence Intervals shown as error bars",
    x = "Beta Coefficient (log-odds scale)",
    y = "Feature",
    fill = "Direction"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "bottom"
  )

print(coef_plot)

# Save plot
ggsave("top10_betas.png", coef_plot, width = 10, height = 7, dpi = 300)
```

# 8. Model Performance: ROC-AUC

```{r roc-auc, echo=TRUE}
# Predictions
predicted_probs <- predict(final_model, type = "response")
predicted_class <- ifelse(predicted_probs > 0.5, 1, 0)

# Confusion Matrix
conf_matrix <- table(Actual = y, Predicted = predicted_class)
cat("Confusion Matrix:\n")
print(conf_matrix)

accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
sensitivity <- conf_matrix[2, 2] / sum(conf_matrix[2, ])
specificity <- conf_matrix[1, 1] / sum(conf_matrix[1, ])

cat("\nPerformance Metrics:\n")
cat("Accuracy:", round(accuracy, 4), "\n")
cat("Sensitivity (Recall):", round(sensitivity, 4), "\n")
cat("Specificity:", round(specificity, 4), "\n")

# ROC Curve
roc_obj <- roc(y, predicted_probs, quiet = TRUE)
auc_value <- auc(roc_obj)

cat("\nROC-AUC:", round(as.numeric(auc_value), 4), "\n")

# Plot ROC Curve
plot(roc_obj, main = "ROC Curve", print.auc = TRUE, 
     auc.polygon = TRUE, grid = TRUE)
```



# 9. Boxplots for Important Predictors (Rescaled Data)

```{r important-predictors-boxplots, echo=TRUE, fig.width=12, fig.height=6}
# Boxplots for the most important predictors from logistic regression
# Uses rescaled data (X_final_rescaled) that were also used for the model

# Define important predictors in order of importance
important_predictors_requested <- c("texture_worst", "concave.points_worst", "compactness_se")

# Check if X_final_rescaled is available, if not, use X_final
if(!exists("X_final_rescaled")) {
  X_final_rescaled <- X_final
  cat("\nNote: X_final_rescaled not available, using X_final\n")
}

# Find correct variable names (may have dot or space)
# Check which variables are actually present
available_predictors <- c()
for(pred in important_predictors_requested) {
  if(pred %in% colnames(X_final_rescaled)) {
    available_predictors <- c(available_predictors, pred)
  } else if(pred == "concave.points_worst") {
    # Try variant with space
    if("concave points_worst" %in% colnames(X_final_rescaled)) {
      available_predictors <- c(available_predictors, "concave points_worst")
    }
  }
}

# Ensure predictors are in the specified order
ordered_predictors <- c()
for(pred in important_predictors_requested) {
  if(pred %in% available_predictors) {
    ordered_predictors <- c(ordered_predictors, pred)
  } else if(pred == "concave.points_worst" && "concave points_worst" %in% available_predictors) {
    ordered_predictors <- c(ordered_predictors, "concave points_worst")
  }
}
available_predictors <- ordered_predictors

if(length(available_predictors) > 0) {
  cat("\nCreating boxplots for important predictors:\n")
  cat(paste(available_predictors, collapse = ", "), "\n")
  
  # Create diagnosis labels
  diagnosis_labels <- ifelse(data$diagnosis == "M", "Malignant", "Benign")
  
  # Create data frame with rescaled data
  plot_data_important <- data.frame(
    diagnosis = diagnosis_labels,
    X_final_rescaled[, available_predictors, drop = FALSE]
  )
  
  # Pivot to long format for facet_wrap
  # Use all columns except diagnosis (robust to variable names)
  plot_data_important_long <- plot_data_important %>%
    pivot_longer(cols = -diagnosis, 
                 names_to = "predictor", 
                 values_to = "value")
  
  # Clean predictor names: convert to character, remove whitespace, standardize naming
  plot_data_important_long$predictor <- as.character(plot_data_important_long$predictor)
  plot_data_important_long$predictor <- trimws(plot_data_important_long$predictor)
  # Replace "concave points_worst" (with space) with "concave.points_worst" (with dot)
  plot_data_important_long$predictor[plot_data_important_long$predictor == "concave points_worst"] <- "concave.points_worst"
  
  # Set factor levels explicitly to ensure correct facet ordering and no NAs
  ordered_predictors <- c("texture_worst", "concave.points_worst", "compactness_se")
  plot_data_important_long$predictor <- factor(plot_data_important_long$predictor, 
                                                levels = ordered_predictors)
  
  # Create boxplots with facet_wrap
  important_boxplots <- ggplot(plot_data_important_long, 
                                aes(x = diagnosis, y = value, fill = diagnosis)) +
    geom_boxplot(alpha = 0.7, outlier.shape = NA) +
    geom_jitter(width = 0.2, alpha = 0.3, size = 0.5) +
    facet_wrap(~ predictor, scales = "free_y", ncol = 3) +
    scale_fill_manual(values = c("Benign" = "#2E8B57", "Malignant" = "#CD5C5C"),
                      name = "Diagnosis") +
    labs(
      title = "Boxplots: Important Predictors by Diagnosis (Benign vs. Malignant)",
      subtitle = paste("Predictors:", paste(available_predictors, collapse = ", ")),
      x = "Diagnosis",
      y = "Feature value"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 9),
      strip.text = element_text(size = 9, face = "bold"),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    )
  
  print(important_boxplots)
  
  # Save plot
  ggsave("important_predictors_boxplots.png", important_boxplots, 
         width = 12, height = 6, dpi = 300)
  cat("\nBoxplots saved as: important_predictors_boxplots.png\n")
} else {
  cat("\nWarning: None of the requested predictors found.\n")
  cat("Requested predictors:", paste(important_predictors_requested, collapse = ", "), "\n")
  cat("Available variables in X_final_rescaled (first 10):\n")
  cat(paste(head(colnames(X_final_rescaled), 10), collapse = ", "), "\n")
}
```

